### 2025-02-09 products/manuscript/Manuscript.qmd
UpdateCode

---
title: "Example Manuscript Template for a Data Analysis Project"
author: "Asmith Joseph"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

#1.Load Required Libraries
install.packages("janitor")      # For cleaning column names
install.packages("lubridate")    # For working with dates

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

```{r}
# Load necessary libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(readr)
library(here)

# Print the root directory to confirm the correct path
print(here())

# Read the CSV file using here()
PROMO_Data <- read_csv(here("PROMO_Data.csv"))

# View the dataset
View(PROMO_Data)

# View the data
View(PROMO_Data)

#3. Initial Data Inspection
# View structure of the dataset
str(PROMO_Data)

# View first few rows
head(PROMO_Data)

# Check column names
colnames(PROMO_Data)

# Get summary statistics
summary(PROMO_Data)

# Count missing values in each column
colSums(is.na(PROMO_Data))

```

**Authors**

* Asmith Joseph$^{1,2,*}$ (ORCID: 0009-0004-6875-0868)

**Author affiliations**

1. College of Public Health, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: asmith.joseph@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
The data for this analysis comes from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, published in The Lancet (Kajubi et al., 2019). This was a double-blind, randomized, controlled, superiority trial conducted to compare monthly IPTp-SP vs. IPTp-DP among pregnant women in malaria-endemic regions.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._
 Does Malaria Preventive Treatment (DP vs. SP) Influence the Likelihood of Recurrent Placental Malaria in Subsequent Pregnancies?
Hypothesis: Women who received DP during pregnancy may have a lower likelihood of recurrent placental malaria in subsequent pregnancies compared to those who received SP, due to better efficacy of DP in preventing malaria infection.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
_The dataset used in this analysis originates from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, a double-blind, randomized controlled trial investigating the effectiveness of monthly intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women and their infants. The trial compared sulfadoxine-pyrimethamine (SP) vs. dihydroartemisinin-piperaquine (DP) in malaria prevention.

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._
-A subset of variables was created from the ClinEpiDB Data Repository, including data from 782 participants. The dataset consists of 138 variables covering various aspects such as:
Study Arm, Participant Demographics, Clinical and Medical History, Obstetric History, Diagnosis and Obstetric Data (Pregnancy, Labor & Delivery, Delivery Complications), Medications Administered During Labor and Delivery, Aggregated Illness Data, Aggregated Lab Tests, Child Demographics
The dataset was downloaded as a text document and processed in RStudio to ensure it was clean and ready for analysis.

The cleaning process involved: Handling Missing Values, Renaming Columns for Consistency, Converting Data Types (e.g., categorical, numerical, date formats), and Removing Duplicate Records.

## Statistical analysis
_Explain anything related to your statistical analyses._
Planned Analysis 
Outcome & Predictors of Interest
Primary Outcome: Placental Malaria (Yes/No)
Key Predictor: Malaria Preventive Treatment (DP vs. SP)
Additional Factors to Consider:
Prior Malaria Exposure (Total malaria episodes before/during pregnancy)
Maternal Characteristics (Gravidity, Parity, Age)
Other Health Conditions (Hypertension, Anemia)

Statistical Plan
Descriptive Statistics: Compare PM rates across DP vs. SP groups.
Logistic Regression: Model the effect of DP vs. SP on the likelihood of PM, adjusting for potential confounders.
Interaction Analysis: Check if prior malaria exposure modifies the effect of DP vs. SP on PM risk.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References